---
title: "tidying"
author: "gokul"
date: "2024-05-03"
output: html_document
---

```{r}
# Attach packages
library(tidyverse)
library(janitor)
library(here)
library(readxl)
install.packages("janitor")
```
library(janitor)

```{r}
inverts <- read_excel(here("data", "inverts.xlsx"))
```
library(readxl)

inverts <- read_excel(here("data", "inverts.xlsx"))

library(here)
inverts <- read_excel(here("data", "inverts.xlsx"))


```{r}
view(inverts)
names(inverts)
summary(inverts)
```
view(inverts)


```{r}
inverts_long <- pivot_longer(data = inverts, cols = "2016" : "2018", names_to = "year", values_to = "sp_count")

inverts_long
```
view(inverts_long)
view(inverts)

```{r}
class(inverts_long$year)
```
```{r}
inverts_long <- inverts_long %>%
  mutate(year = as.numeric(year))


inverts_wide <- inverts_long %>%
  pivot_wider(names_from = common_name, values_from = sp_count)


inverts_wide
view(inverts_wide)

inverts_wide <- inverts_wide %>%
  clean_names()

inverts_unite <- inverts_long %>%
  unite(col = "site_year", c(site,year), sep = "_")


view(inverts_unite)


inverts_moyr <- inverts_long %>%
  unite(col = "mo_yr", c(month, year), sep = " hello ")

view(inverts_moyr)




inverts_sep <- inverts_unite %>%
  tidyr::separate(site_year, into = c("my _site","my_year"))
view(inverts_sep)


ca_abbr <- inverts %>%
  mutate( common_name = str_replace(common_name, pattern = "california",replacement = "CA"))
view(ca_abbr)
```

